Seniority and hierarchy configuration interaction for radicals and excited states

### 2023-08-18 Manuscript/manuscript.tex
abstract

% Abstract
\begin{abstract}
{\bf Abstract:}
Hierarchy configuration interaction (hCI) has been recently introduced as an alternative configuration interaction (CI) route combining excitation degree and seniority number,
which showed to efficiently recover dynamic and static correlation for closed-shell molecular systems [J. Phys. Chem. Lett. 2022, 13, 19, 4342].
Here, we generalize hCI for arbitrary reference determinants, allowing calculations for radicals and for excited states in a state-specific approach.
We gauge this route against excitation-based CI (eCI) and seniority-based CI (sCI)
by evaluating how different properties of ground-state radicals converge to the full CI limit.
hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
Employing the second-order Epstein-Nesbet perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
Our results underline that both the choice of reference and orbitals drive the fine balance between correlation of ground and excited states.
State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results.
In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled cluster methods.
%\bigskip
%\begin{center}
%        \boxed{\includegraphics[keepaspectratio,width=2in]{TOC}}
%\end{center}
%\bigskip
\end{abstract}

>>>>

% Abstract
\begin{abstract}
Hierarchy configuration interaction (hCI) has been recently introduced as an alternative configuration interaction (CI) route combining excitation degree and seniority number, which showed to efficiently recover both dynamic and static correlations for closed-shell molecular systems [\href{https://doi.org/10.1021/acs.jpclett.2c00730}{\textit{J.~Phys.~Chem.~Lett.}~\textbf{2022}, \textit{13}, 4342}].
Here, we generalize hCI for an arbitrary reference determinant, allowing calculations for radicals and for excited states in a state-specific way.
We gauge this route against excitation-based CI (eCI) and seniority-based CI (sCI) by evaluating how different ground-state properties of radicals converge to the full CI limit.
We find that hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
Employing the second-order Epstein-Nesbet perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
Our results underline that both the choice of the reference determinant and the set of orbitals drive the fine balance \titou{between correlation of ground and excited states}.
State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results.
In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled cluster methods.
%\bigskip
%\begin{center}
%        \boxed{\includegraphics[keepaspectratio,width=2in]{TOC}}
%\end{center}
%\bigskip
\end{abstract}
